---
title: 'Assignment 2: Data visualization'
author: "Tamas Nagy"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytuesdayR)
library(rmarkdown)
library(viridis)
library(ggplot2)
library(RColorBrewer)
library(scales)
```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r}
#
climbing_data <- tt_load("2020-09-22")
expeditions <- climbing_data$expeditions
head(expeditions)
```

```{r}
expeditions2 <- expeditions %>%
  mutate(peak_name = fct_lump(peak_name, n = 15)) %>%
  filter(peak_name != "Other") %>%
  group_by(peak_name, season) %>%
  summarise(expedition_number = n(), .groups = 'drop') %>%
  mutate(peak_name = fct_reorder(peak_name, expedition_number, .fun = sum))

str(expeditions2)
#print(head(expedition2))
```

```{r}
task1 <- ggplot(expeditions2, aes(x = expedition_number, y = peak_name, fill = season)) +
  geom_col() +
  scale_fill_viridis(discrete = TRUE) +
  labs(title = "The 15 most popular peaks stacked by season of expedition", x = "Number of Expeditions") +
  theme_light() +
  theme(legend.position = "bottom", legend.direction = "horizontal")

print(task1)
```

## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r}
phd_by_field <- tt_load("2019-02-19")
phd_data <- phd_by_field$phd_by_field
head(phd_data)
```

```{r}
phd_summary <- phd_data %>%
  group_by(broad_field, year) %>%
  summarise(total_phds = sum(n_phds, na.rm = TRUE))

print(phd_summary)
```

```{r}

task2 <- ggplot(phd_summary, aes(x = year, y = total_phds, color = broad_field)) +
  geom_line(linewidth = 1.2) +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = pretty_breaks(n = 5)) +
  scale_y_continuous(labels = comma_format()) + 
  labs(title = "Number of awarded Ph.D.'s in the US by year", x = NULL, y = NULL, color = "Broad field") +
  theme_minimal()

print(task2)

```

## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.

```{r}
commute <- tt_load("2019-11-05")
commute_data <- commute$commute
head(commute_data)
```

```{r}
commute_summary <- commute_data %>%
  group_by(state_abb, mode, state_region) %>%
  summarise(N = sum(n))
  
  
print(commute_summary)
```

```{r}
commute_data_wide <- commute_summary %>%
  pivot_wider(names_from = mode, values_from = N)

print(commute_data_wide)
```

```{r}
task3 <- ggplot(commute_data_wide, aes(x = Walk, y = Bike, label = state_abb)) +
  geom_point(aes(color = state_region)) +
  geom_text(vjust = 0.5, hjust = 0.5, check_overlap = TRUE) +
  scale_x_log10(labels = comma_format()) +
  scale_y_log10(labels = comma_format(), limits = c(20, NA)) +
  labs(title = "Title number pf people walking vs biking to work in each USA state",
       x = "Number of ppl walking to work (log N)",
       y = "Number of ppl bikinng to work (log N)",
      color = "State region") +
  theme_light()
  theme(legend.position = "right", name = "State region")
  
print(task3)
```
